Czech parliament meeting recordings as ASR training data

### 2020-07-01 2020-fedcsis-sofia-svolocz/svolocz.tex
fedcsis svolocz: numerals

>>>>

There are many numeral expressions in the transcripts. They amount to 489,880
out of 25,010,269 tokens in the complete stenographic transcript, which is
almost two percent. In the training dataset, 24\% of the samples contain one or
more numerals.

Originally, I have included the digits into the alphabet for speech recognition,
thus attempting to train the system to transcribe numeral expressions directly
into digits. The speech recognition system described in the following section
would however transcribe numeral expressions as empty strings.

There are four ways to deal with the problem:
\begin{enumerate}
\item{ignore it,}
\item{remove digits from the training data,}
\item{manually expand digits to words,}
\item{automatically expand digits to words.}
\end{enumerate}

The first option needs no elaboration. The second one, removing samples with
digits, is an easy and viable option but it is a waste of a quarter of the
dataset and of the vast majority of samples with numerals in them. Manual
expansion would surely be ideal but very costly. It remains to attempt the
fourth variant of automated expansion.

For automated expansion of digits into words, we can use the available initial
transcript and the algorithm for alignment with the stenographic transcript.

The expansion is done in two steps:
\begin{enumerate}
\item{generation of verbal variants,}
\item{selection of the most likely variant.}
\end{enumerate}

I have used the Perl module \texttt{Lingua::CS::Num2Word} as a base for the
expansion. I modified the module in the following way: 1) I added support
the order of billions, which is very common in the corpus. 2) A number is no
longer expanded into a single phrase but instead into all possible phrases
expressing the given number. 3) I added support for genitive and accusative
cases, decimal numerals, ordinals, dates and times.

All tokens in the stenographs that include digits are expanded into their
verbalization variants before further processing. Upon alignment, the variant
with least edit distance from the initial transcript is selected.

Common abbreviations and symbols are expanded together with the digits. For
example, the very common character \textit{``§'' (paragraph)} is expanded into
the forms \textit{paragraf, paragrafu, paragrafů, paragrafem,
paragrafech} that represent common inflections of the word. Some common
abbreviations that undergo inflection include \textit{``čl.'' (article)},
\textit{``odst.''}, which also means ``paragraph'' and \textit{``tzv.''
(co-called)}.

With the expansion in the process, the similarity of the stenographic transcript
and the initial one raised, which raised the number of accepted segments from
26\% to 35\%. The amount of training data grew by 86 hours to 1144.
